# Data
For this research/analysis, we will mainly use the Residential Energy Consumption Survey (RECS) 2020 data produced by the U.S Energy Information Administration (EIA). The 2020 RECS study represents the 15th iteration of the program, which started in 2020 and the last update was completed in Jan 2020. Data was collected from around 18,500 households in housing units statistically selected to represent the 123.5 million housing units that are occupied as primary residences (EIA, 2024). The survey offers data in:

- Detailed information on energy usage patterns in U.S. homes
- Data on various energy end uses, including air conditioning, space heating, and water heating
- Breakdowns of electricity consumption by region and housing type

## Description
There is a total of 799 features in this dataset, but for the purpose of this analysis, we investigate the relationship between household total electricity use (KWH) with the following features:

1. Climatic Impact:
    1. HDD65 (Heating degree days in 2020)
    2. CDD65 (Cooling degree days in 2020)
    3. BA_climate (Building America Climate Zone)
2. Geographic Impact:
    1. state_postal (State and State Postal)
    2. UATYP10 (2010 Census Urban Type Code)
    3. TYPEHUQ (Type of housing unit)
3. Socioeconomic Factors
    1. MONEYPY (Annual gross household income for the past year)
    2. ENERGYASST20 (Received energy assistance)
4. Habitual factors
    1. COOLCNTL (Summer temperature control method)
    2. LGTINLED (Portion of inside light bulbs that are LED)
    
As these data are all unified under the same source, we can analyze from various economic and geographic dimensions without the need to merge many datasets and deal with the different granularity of data. 

## Missing value analysis
```{r, message = FALSE, warning=FALSE}
# Setup
library(tidyverse)
df <- read.csv('recs2020.csv')

# We want to understand which columns have the most na's
# This also helps us know if the columns we are using are affected
na_counts <- colSums(is.na(df)) |> sort(decreasing = TRUE)  
# head(na_counts, 15)

# Select only the columns we are interested in
energy <- df |>
  select(state_name, state_postal, KWH, HDD65, CDD65, UATYP10, TYPEHUQ, MONEYPY, ENERGYASST20, BA_climate, COOLCNTL, HEATCNTL, LGTINLED)
```

We first understand that the columns with missing data are not in the set of columns that we need to use. However, we also want to check if the KWH values are 0 (which may indicate missing data). The check below shows that no KWH values are 0 as well, indicating that there is no potential missing or problematic data within the part of the dataset that we are working with,

```{r, message = FALSE, warning=FALSE}

# Check KWH column if anything equals to 0
sum(energy$KWH == 0)
library(redav)
plot_missing(energy, percent = FALSE)
```

This check suggests that the section of the dataset we are using for this project is complete and we do not need to work with missing data. However, closer inspection of the specific columns suggests numerical values that represent missing values. These are in general cleaned and usually do not have a high proportion within the dataset. Hence, we will remove these na values when necessary during the latter parts of the analysis.


## Annex A
This section can be used to refer to self-defined values for various features in the results section.

COOLCNTL
1: (Set and leave it) Set one temperature and leave it there most of the time
2: (Manual) Manually adjust the temperature 
3: (Autoset) Programmable or smart thermostat automatically adjusts the temperature 
4: (Manual) Turn equipment on or off as needed
5: (No control) Our household does not have control over the temperature

MONEYPY (Income band)
Based on the average household income in the selected states, we simplify the categories to the below:
Below 35,000 set to low (categories 1-9)
From 35,000-100,000 set to mid (categories 10-14)
Higher than 100,000 set to upper (categories 14-16)


KWH (Energy usage) 
Set <25% to 0-9300 kwh, 
Set 25-50% to 9300-13400, 
Set 50-75% to 13400-18400, 
Set >75% to above 18400

COOLCNTL (Cooling )
1: Set one temperature and leave it there most of the time (Set and leave)
2: Manually adjust the temperature (Manual)
3: Programmable or smart thermostat automatically adjusts the temperature (Autoset)
4: Turn equipment on or off as needed (Manual)
5: No control (no control)
Category 2 and 4 incurs lower energy usage and are recommended

