# Data
For this research/analysis, we will mainly use the Residential Energy Consumption Survey (RECS) 2020 data produced by the U.S Energy Information Administration (EIA). The 2020 RECS study represents the 15th iteration of the program, which started in 2020 and the last update was completed in Jan 2020. Data was collected from around 18,500 households in housing units statistically selected to represent the 123.5 million housing units that are occupied as primary residences (EIA, 2024). The survey offers data in:

- Detailed information on energy usage patterns in U.S. homes
- Data on various energy end uses, including air conditioning, space heating, and water heating
- Breakdowns of electricity consumption by region and housing type

## Description
There is a total of 799 features in this dataset, but for the purpose of this analysis, we investigate the relationship between household total electricity use (KWH) with the following features:

1. Climatic Impact:
    1. HDD65 (Heating degree days in 2020)
    2. CDD65 (Cooling degree days in 2020)
2. Geographic Impact:
    1. state_postal (State and State Postal)
    2. UATYP10 (2010 Census Urban Type Code)
    3. TYPEHUQ (Type of housing unit)
3. Socioeconomic Factors
    1. MONEYPY (Annual gross household income for the past year)
    2. ENERGYASST20 (Received energy assistance )
  
As these data are all unified under the same source, we can analyze from various economic and geographic dimensions without the need to merge many datasets and deal with the different granularity of data. Hence, this study mainly uses the RECS dataset, but may include other data sources later in the project based on necessity.

## Missing value analysis
```{r, message = FALSE, warning=FALSE}
# Setup
library(tidyverse)
df <- read.csv('recs2020.csv')

# We want to understand which columns have the most na's
# This also helps us know if the columns we are using are affected
na_counts <- colSums(is.na(df)) |> sort(decreasing = TRUE)  
# head(na_counts, 15)

# Select only the columns we are interested in
energy <- df |>
  select(state_postal, KWH, DOLLAREL, HDD65, CDD65, UATYP10, TYPEHUQ, MONEYPY, ENERGYASST20)
```

We first understand that the columns with missing data are not in the set of columns that we need to use. However, we also want to check if the KWH values are 0 (which may indicate missing data). The check below shows that no KWH values are 0 as well, indicating that there is no potential missing or problematic data within the part of the dataset that we are working with,

```{r, message = FALSE, warning=FALSE}

# Check KWH column if anything equals to 0
sum(energy$KWH == 0)
library(redav)
plot_missing(energy, percent = FALSE)
```
This check suggests that the section of the dataset we are using for this project is complete and we do not need to work with missing data.

